Master thesis

### 2017-06-09 03-Model.Rmd
Appendix 3

20 to 30 % of designated trees are considered as rotten once probed by the lumberman.
Rotten trees are not random and depends both on tree species and diameter.

**Explain rotten bayesian model**

\begin{equation}
  P_{rotten} = logit^{-1}(\beta_0 + \beta_1*dbh) = \frac{e^{\beta_0 + \beta_1*dbh}}{1 + e^{\beta_0 + \beta_1*dbh}}
  (\#eq:rotten)
\end{equation}

(see [Appendix 3: Rotten tree model] for more details)

@@

Most of models account long term damages due to selective logging with a 10 years increased mortality [@Huth2004; @Khler2004; @Ruger2008]. We decided to model explicitly long term logging damages because of their localised nature through a gap damages model. We gathered data from Paracou dataset **Cite** in cenususes between 1988 and 1992 on Paracou harvested plots. Individuals were categorized between alive, dead, or recruited during the period. We measured each individual distance to the closest gap. We then used a bayesian approach to test the link between tree death in the four years following the log event and distance to the closest gap. We adapted the model from **Cite Herault** based on a disturbance index into.

\begin{equation}
  \begin{array}{c} 
    Death_i \sim \mathcal{B}(P(Death_i)) \\
    P(Death_i) = logit^{-1}(\theta + \beta*e^{\alpha*d_{gaps}}) = \frac{e^{\theta + \beta*e^{\alpha*d_{gaps}}}}{1 + e^{\theta + \beta*e^{\alpha*d_{gaps}}}}
  \end{array}
  (\#eq:death)
\end{equation}

$Death$ of a tree $i$ follows a $\mathcal{Bernoulli}$ law of probability $P(Death_i)$. The odds for a tree $i$ to die $O(Death_i)$ is calculated by the sum of the natural tree death odd $\theta$ and a perturbation index $\beta*e^{\alpha*d_{gaps}}$. The perturbation index depend on the distance $d_{gaps}$ of the tree $i$ to the closest logging gap. The probability for a tree $i$ to die $P(Death_i)$ is finally calculated by taking the inverse logit of the odd $logit^{-1}[O(Death_i)]$ (see [Appendix 4: Gap damages model] for more details).

>>>>

20 to 30 % of designated trees are considered as rotten once probed by the lumberman, and thus not harvested. Rotten trees are not random and depends both on tree species and diameter. We gathered data from the forest office (ONF, Laurent Descroix, personnal communication) inventories precising if tree were probbed as rotten and their corresponding species and diameter. In addition, tree plots and sawed volume was informed. We then used a bayesian approach to model the link between tree species and diameter and their risk to be probbed as rotten by the lumberman. We test different models with growing level of complexity and kept the model with the best tradeoff between complexity (number of parameters), convergence, likelihood, and prediction quality (root mean square error of prediction RMSEP):

\begin{equation}
  \begin{array}{c} 
    probbed~rotten \sim \mathcal{B}(P(probbed~rotten)) \\
    P(probbed~rotten) = logit^{-1}(\beta_0 + \beta_1*dbh) = \frac{e^{\beta_0 + \beta_1*dbh}}{1 + e^{\beta_0 + \beta_1*dbh}}
  \end{array}
  (\#eq:rotten)
\end{equation}

Tree $probbed~rotten$ follows a $Bernoulli$ law of probability $P(probbed~rotten)$. The odds for a tree to be probbed as rotten are calculated with the sum of a base odd to be rotten $\beta_0$ and a diameter dependent odd calculated with $\beta_1$. The probability for a tree to be probbed as rotten $P(probbed~rotten)$ is finally calculated by taking the inverse logit $logit^{-1}$ of the odd (see [Appendix 3: Rotten tree model] for more details).

@@

Most of models account long term damages due to selective logging with a 10 years increased mortality [@Huth2004; @Khler2004; @Ruger2008]. We decided to model explicitly long term logging damages because of their localised nature through a gap damages model. We gathered data from Paracou dataset [@Guehl2004] in cenususes between 1988 and 1992 on Paracou harvested plots. Individuals were categorized between alive, dead, or recruited during the period. We measured each individual distance to the closest gap. We then used a bayesian approach to test the link between tree death in the four years following the log event and distance to the closest gap. We adapted the model from @Herault2010 based on a disturbance index into:

\begin{equation}
  \begin{array}{c} 
    Death \sim \mathcal{B}(P(Death)) \\
    P(Death) = logit^{-1}(\theta + \beta*e^{\alpha*d_{gaps}}) = \frac{e^{\theta + \beta*e^{\alpha*d_{gaps}}}}{1 + e^{\theta + \beta*e^{\alpha*d_{gaps}}}}
  \end{array}
  (\#eq:death)
\end{equation}

$Death$ of a tree follows a $\mathcal{Bernoulli}$ law of probability $P(Death)$. The odds for a tree to die are calculated with the sum of the natural tree death odd $\theta$ and a perturbation index $\beta*e^{\alpha*d_{gaps}}$. The perturbation index depend on the distance $d_{gaps}$ of the tree $i$ to the closest logging gap. The probability for a tree to die $P(Death)$ is finally calculated by taking the inverse logit $logit^{-1}$ of the odd.

### 2017-06-09 08-Appendix2.Rmd
Appendix 3

```{r A2traits, echo=FALSE}
knitr::kable(data.frame(
  Name = c('LL', 'SLA', 'N', 'P', 'wsg'),
  Trait = c('Leaf lifespan (longevity)',
            'Leaf area per leaf dry mass (specific leaf area, SLA)',
            'Leaf nitrogen (N) content per leaf dry mass',
            'Leaf phosphorus (P) content per leaf dry mass',
            'Stem dry mass per stem fresh volume (stem specific density)'),
  Unit = c('month',' $m^2.kg^{-1}$', '$mg.g^{-1}$', '$mg.g^{-1}$', '$mg.mm^{-3}$'),
  TRYcode = c(12, 11, 15, 14, 4)),
  caption = 'Functional traits gathered with TRY.')
```

@@

```{r A2LLpred, echo=FALSE, fig.cap='Leaflifespan predictions for the selected model with leaf mass per area (LMA), leaf nitrogen content (Nmass), wood specific gravity (wsg) and predicted versus observed values. Leaf lifespan (LL) is predicted with model M10 fit. Leaf mass per area (LMA) and leaf nitrogen content (Nmass), and wood specific gravity (wsg) are taken in a composite dataset of GLOPNET, TRY and DRYAD  datasets.Warning LMA (resp. Nmass and wsg) is not constant and depend on the closest point value for right (resp. center and left) graph.', fig.height=12, fig.width=12}

>>>>

```{r A2traits, echo=FALSE}
knitr::kable(data.frame(
  Name = c('LL', 'SLA', 'N', 'P', 'wsg'),
  Trait = c('Leaf lifespan (longevity)',
            'Leaf area per leaf dry mass (specific leaf area, SLA)',
            'Leaf nitrogen (N) content per leaf dry mass',
            'Leaf phosphorus (P) content per leaf dry mass',
            'Stem dry mass per stem fresh volume (stem specific density)'),
  Unit = c('month',' $m^2.kg^{-1}$', '$mg.g^{-1}$', '$mg.g^{-1}$', '$mg.mm^{-3}$'),
  TRYcode = c(12, 11, 15, 14, 4)),
  caption = 'Functional traits gathered with TRY.', format = 'pandoc')
```

@@

```{r A2LLpred, echo=FALSE, fig.cap='Leaflifespan predictions for the selected model with leaf mass per area (LMA), leaf nitrogen content (Nmass), wood specific gravity (wsg) and predicted versus observed values. Leaf lifespan (LL) is predicted with model M10 fit. Leaf mass per area (LMA) and leaf nitrogen content (Nmass), and wood specific gravity (wsg) are taken in a composite dataset of GLOPNET, TRY and DRYAD  datasets.Warning LMA (resp. Nmass and wsg) is not constant and depend on the closest point value for right (resp. center and left) graph.', cahce=TRUE, fig.height=12, fig.width=12}

### 2017-06-09 docs/master-thesis.tex
Appendix 3

20 to 30 \% of designated trees are considered as rotten once probed by
the lumberman. Rotten trees are not random and depends both on tree
species and diameter.

\textbf{Explain rotten bayesian model}

\begin{equation}
  P_{rotten} = logit^{-1}(\beta_0 + \beta_1*dbh) = \frac{e^{\beta_0 + \beta_1*dbh}}{1 + e^{\beta_0 + \beta_1*dbh}}
  \label{eq:rotten}
\end{equation}

(see \protect\hyperlink{appendix-3-rotten-tree-model}{Appendix 3: Rotten
tree model} for more details)

@@

Most of models account long term damages due to selective logging with a
10 years increased mortality \citep{Huth2004, Khler2004, Ruger2008}. We
decided to model explicitly long term logging damages because of their
localised nature through a gap damages model. We gathered data from
Paracou dataset \textbf{Cite} in cenususes between 1988 and 1992 on
Paracou harvested plots. Individuals were categorized between alive,
dead, or recruited during the period. We measured each individual
distance to the closest gap. We then used a bayesian approach to test
the link between tree death in the four years following the log event
and distance to the closest gap. We adapted the model from \textbf{Cite
Herault} based on a disturbance index into.

\begin{equation}
  \begin{array}{c} 
    Death_i \sim \mathcal{B}(P(Death_i)) \\
    P(Death_i) = logit^{-1}(\theta + \beta*e^{\alpha*d_{gaps}}) = \frac{e^{\theta + \beta*e^{\alpha*d_{gaps}}}}{1 + e^{\theta + \beta*e^{\alpha*d_{gaps}}}}
  \end{array}
  \label{eq:death}
\end{equation}

\(Death\) of a tree \(i\) follows a \(\mathcal{Bernoulli}\) law of
probability \(P(Death_i)\). The odds for a tree \(i\) to die
\(O(Death_i)\) is calculated by the sum of the natural tree death odd
\(\theta\) and a perturbation index \(\beta*e^{\alpha*d_{gaps}}\). The
perturbation index depend on the distance \(d_{gaps}\) of the tree \(i\)
to the closest logging gap. The probability for a tree \(i\) to die
\(P(Death_i)\) is finally calculated by taking the inverse logit of the
odd \(logit^{-1}[O(Death_i)]\) (see
\protect\hyperlink{appendix-4-gap-damages-model}{Appendix 4: Gap damages
model} for more details).

@@

\begin{table}

\caption{\label{tab:A2traits}Functional traits gathered with TRY.}
\centering
\begin{tabular}[t]{l|l|l|r}
\hline
Name & Trait & Unit & TRYcode\\
\hline
LL & Leaf lifespan (longevity) & month & 12\\
\hline
SLA & Leaf area per leaf dry mass (specific leaf area, SLA) & \$m\textasciicircum{}2.kg\textasciicircum{}\{-1\}\$ & 11\\
\hline
N & Leaf nitrogen (N) content per leaf dry mass & \$mg.g\textasciicircum{}\{-1\}\$ & 15\\
\hline
P & Leaf phosphorus (P) content per leaf dry mass & \$mg.g\textasciicircum{}\{-1\}\$ & 14\\
\hline
wsg & Stem dry mass per stem fresh volume (stem specific density) & \$mg.mm\textasciicircum{}\{-3\}\$ & 4\\
\hline
\end{tabular}
\end{table}

@@

\hypertarget{appendix-4-gap-damages-model}{\section{Appendix 4: Gap
damages model}\label{appendix-4-gap-damages-model}}

>>>>

20 to 30 \% of designated trees are considered as rotten once probed by
the lumberman, and thus not harvested. Rotten trees are not random and
depends both on tree species and diameter. We gathered data from the
forest office (ONF, Laurent Descroix, personnal communication)
inventories precising if tree were probbed as rotten and their
corresponding species and diameter. In addition, tree plots and sawed
volume was informed. We then used a bayesian approach to model the link
between tree species and diameter and their risk to be probbed as rotten
by the lumberman. We test different models with growing level of
complexity and kept the model with the best tradeoff between complexity
(number of parameters), convergence, likelihood, and prediction quality
(root mean square error of prediction RMSEP):

\begin{equation}
  \begin{array}{c} 
    probbed~rotten \sim \mathcal{B}(P(probbed~rotten)) \\
    P(probbed~rotten) = logit^{-1}(\beta_0 + \beta_1*dbh) = \frac{e^{\beta_0 + \beta_1*dbh}}{1 + e^{\beta_0 + \beta_1*dbh}}
  \end{array}
  \label{eq:rotten}
\end{equation}

Tree \(probbed~rotten\) follows a \(Bernoulli\) law of probability
\(P(probbed~rotten)\). The odds for a tree to be probbed as rotten are
calculated with the sum of a base odd to be rotten \(\beta_0\) and a
diameter dependent odd calculated with \(\beta_1\). The probability for
a tree to be probbed as rotten \(P(probbed~rotten)\) is finally
calculated by taking the inverse logit \(logit^{-1}\) of the odd (see
\protect\hyperlink{appendix-3-rotten-tree-model}{Appendix 3: Rotten tree
model} for more details).

@@

Most of models account long term damages due to selective logging with a
10 years increased mortality \citep{Huth2004, Khler2004, Ruger2008}. We
decided to model explicitly long term logging damages because of their
localised nature through a gap damages model. We gathered data from
Paracou dataset \citep{Guehl2004} in cenususes between 1988 and 1992 on
Paracou harvested plots. Individuals were categorized between alive,
dead, or recruited during the period. We measured each individual
distance to the closest gap. We then used a bayesian approach to test
the link between tree death in the four years following the log event
and distance to the closest gap. We adapted the model from
\citet{Herault2010} based on a disturbance index into:

\begin{equation}
  \begin{array}{c} 
    Death \sim \mathcal{B}(P(Death)) \\
    P(Death) = logit^{-1}(\theta + \beta*e^{\alpha*d_{gaps}}) = \frac{e^{\theta + \beta*e^{\alpha*d_{gaps}}}}{1 + e^{\theta + \beta*e^{\alpha*d_{gaps}}}}
  \end{array}
  \label{eq:death}
\end{equation}

\(Death\) of a tree follows a \(\mathcal{Bernoulli}\) law of probability
\(P(Death)\). The odds for a tree to die are calculated with the sum of
the natural tree death odd \(\theta\) and a perturbation index
\(\beta*e^{\alpha*d_{gaps}}\). The perturbation index depend on the
distance \(d_{gaps}\) of the tree \(i\) to the closest logging gap. The
probability for a tree to die \(P(Death)\) is finally calculated by
taking the inverse logit \(logit^{-1}\) of the odd.

@@

\begin{longtable}[]{@{}lllr@{}}
\caption{\label{tab:A2traits}Functional traits gathered with
TRY.}\tabularnewline
\toprule
Name & Trait & Unit & TRYcode\tabularnewline
\midrule
\endfirsthead
\toprule
Name & Trait & Unit & TRYcode\tabularnewline
\midrule
\endhead
LL & Leaf lifespan (longevity) & month & 12\tabularnewline
SLA & Leaf area per leaf dry mass (specific leaf area, SLA) &
\(m^2.kg^{-1}\) & 11\tabularnewline
N & Leaf nitrogen (N) content per leaf dry mass & \(mg.g^{-1}\) &
15\tabularnewline
P & Leaf phosphorus (P) content per leaf dry mass & \(mg.g^{-1}\) &
14\tabularnewline
wsg & Stem dry mass per stem fresh volume (stem specific density) &
\(mg.mm^{-3}\) & 4\tabularnewline
\bottomrule
\end{longtable}

@@

In order to simulate sylviculture with TROLL we need to implement a new
sylviculture module inside TROLL model code. A first litterature review
was completed by an interview with Laurent Descroix of the Office
Nationale des Forêts. We discovered that rotten trees were not random
and seemed to depend both on tree species and diameter. This document
presents modelling of relation between rotten trees and their species
and diameter.

In fact we have two different questions:

\begin{itemize}
\tightlist
\item
  Predict if a tree will be probed as rotten (models \textbf{M})
\item
  Predict how much of tree volume is rotten (models \textbf{N})
\end{itemize}

First all \textbf{M} model can be written as follow:
\[Rotten_n \sim \mathcal{B}(\theta_n), ~~n \in [1,N_{=3816}] ~~ p \in [1, P_{=8}], ~~ s \in [1, S_{=43}]\]
Secondly, all \textbf{N} models depend on a latent variable being the
percentage of rotten wood \(Pt_r\). We can assume that all trees are
growing depending on species \(s\) and plot \(p\) fertility and are
supposed to have a full healthy volume \(V_h\) for a given diameter
\(dbh\). We obtain following model:

\[V_f \sim \mathcal{logN}(V_h*Pt_r, \sigma), ~~n \in [1,N_{=3268}] ~~ p \in [1, P_{=8}], ~~ s \in [1, S_{=43}]\]

We retained following models :

\begin{longtable}[]{@{}ll@{}}
\caption{\label{tab:A3models}Models summary.}\tabularnewline
\toprule
M & Model\tabularnewline
\midrule
\endfirsthead
\toprule
M & Model\tabularnewline
\midrule
\endhead
\(M_{s,p}\) &
\({P_{rotten}}_n \sim \mathcal{B}(inv_{logit}(\beta_0 + \beta_1*dbh_n + {\beta_2}_p + {\beta_3}_s))\)\tabularnewline
\(N_{s,p} + L_{s,p}\) &
\(Volume_{of~wood} \sim \mathcal{logN} (log[((\beta + \beta_p + \beta_s)*dbh^2)*(1 - Pr*((\theta + \theta_p + \theta_s) *dbh^2))], \sigma)\)\tabularnewline
\bottomrule
\end{longtable}

\subsection{Probed rotten (M)}\label{probed-rotten-m}

Based on complexity (number of parameters), convergence and likelihood
we selected model \(M_{p,s}\):

\(M_{s,p}\):
\({P_{rotten}}_n \sim \mathcal{B}(inv_{logit}(\beta_0 + \beta_1*dbh_n + {\beta_2}_p + {\beta_3}_s))\)

\includegraphics{master-thesis_files/figure-latex/A3Mpred-1.pdf}

\begin{longtable}[]{@{}lrrrrrrrrrrrrrrrrr@{}}
\caption{\label{tab:A3Mtab}Models prediction. Probability to be probed as
rotten (P in \%) for a given dbh (cm).}\tabularnewline
\toprule
& 45 & 50 & 55 & 60 & 65 & 70 & 75 & 80 & 85 & 90 & 95 & 100 & 105 & 110
& 115 & 120 & 125\tabularnewline
\midrule
\endfirsthead
\toprule
& 45 & 50 & 55 & 60 & 65 & 70 & 75 & 80 & 85 & 90 & 95 & 100 & 105 & 110
& 115 & 120 & 125\tabularnewline
\midrule
\endhead
P & 4 & 4 & 5 & 7 & 8 & 10 & 12 & 14 & 17 & 20 & 24 & 28 & 32 & 37 & 42
& 47 & 52\tabularnewline
\bottomrule
\end{longtable}

\subsection{Rotten volume (N)}\label{rotten-volume-n}

Based on complexity (number of parameters), convergence and likelihood
we selected model \(N_{p,s}\) associated to hyperparameter \(\rho\) with
model \(L_{p,s}\):

\(N_{s,p} + L_{s,p}\):
\(Volume_{of~wood} \sim \mathcal{logN} (log[((\beta + \beta_p + \beta_s)*dbh^2)*(1 - Pr*((\theta + \theta_p + \theta_s) *dbh^2))], \sigma)\)

\includegraphics{master-thesis_files/figure-latex/A3Npred-1.pdf}

\begin{longtable}[]{@{}lrrrrrrrrrrrrrrrrr@{}}
\caption{\label{tab:A3Ntab}Models prediction. Final volume of wood (\(V_f\)
in \(m^3\)) and percent of rotten wood (\(V_p\) in \%) for a given dbh
(cm) if the tree was probed rotten.}\tabularnewline
\toprule
& 45 & 50 & 55 & 60 & 65 & 70 & 75 & 80 & 85 & 90 & 95 & 100 & 105 & 110
& 115 & 120 & 125\tabularnewline
\midrule
\endfirsthead
\toprule
& 45 & 50 & 55 & 60 & 65 & 70 & 75 & 80 & 85 & 90 & 95 & 100 & 105 & 110
& 115 & 120 & 125\tabularnewline
\midrule
\endhead
Vf & 1.66 & 2.02 & 2.39 & 2.78 & 3.18 & 3.58 & 3.98 & 4.37 & 4.74 & 5.09
& 5.41 & 5.68 & 5.9 & 6.06 & 6.15 & 6.16 & 6.08\tabularnewline
Vp & 7.00 & 9.00 & 11.00 & 13.00 & 15.00 & 18.00 & 20.00 & 23.00 & 26.00
& 29.00 & 32.00 & 36.00 & 40.0 & 43.00 & 47.00 & 52.00 &
56.00\tabularnewline
\bottomrule
\end{longtable}
